---
title: "Challenge 4"
author: "Trenton Tadaki"
format:
  html: 
    self-contained: true
    code-tools: true
    code-fold: true
execute: 
  echo: true
  error: true
  message: false
  warning: false
---

## **Challenge 4**

A joke in the media is that Millennials can't afford houses, because they spend all their money buying avocado toast. Let's use this dataset to address that claim.

> Find or make a dataset with house prices for these four California regions. Join this dataset with your California avocado dataset.

> Use your new joined dataset to make an argument about the relationship between house prices and avocado prices/sales.

> Support your argument with a plot.

```{r Set up}
library(tidyverse)
library(here)
library(lubridate)

housing_prices_og <- read_csv(here("Week 4", "Challenge 4", "City_zhvi_uc_sfrcondo_tier_0.33_0.67_sm_sa_month.csv"))

avocado <- read_csv(here("Week 4", "Lab 4", "avocado.csv"))

cali_regions <- tibble(region = c("LosAngeles", "SanDiego", "Sacramento", "SanFrancisco"))

avocado_clean <- avocado |>
  rename("Small_Medium" = "4046", 
         "Large" = "4225",
         "Extra Large" = "4770")
```

```{r}
avocado_new <- avocado_clean |>
  semi_join(y = cali_regions, by = 'region') |>
  mutate('Avo_sales_USD' = AveragePrice * `Total Volume`,
         'Date' = as.Date(Date),
         'month' = floor_date(Date, 'month')) |>
  select(month, Avo_sales_USD, region) |>
  group_by(region, month) |>
  summarise(Avo_sales_USD = sum(Avo_sales_USD))
  

housing_prices <- housing_prices_og |>
  pivot_longer("2015-01-31" : "2018-03-31",
             names_to = "Date",
             values_to = "Avg_house_price") |>
  filter(State == "CA") |>
  mutate('region' =
           str_replace(RegionName,
                       " ", 
                       ""),
         'Date' = as.Date(Date),
         'month' = floor_date(Date, 'month')) |>
  select(region, month, Avg_house_price) |>
  right_join(cali_regions) |>
  drop_na() 

#REVISION
#Before, I had used multiple mutate functions to replace each of the city names with ones without the spaces. However, now I know that I can simply use the str_replace function to edit the strings directly.

avo_house <- housing_prices |>
  full_join(avocado_new, by = c('region', 'month'))

#REVISION
#Before, I did not have a good grasp on the join functions, so I used a merge() function to join my two dataframes. However, I was required to use full_join() for the purpose of this assignment.


ggplot(data = avo_house) +
  geom_line(mapping = aes(x = month, y = Avg_house_price), color = "blue") +
  geom_line(mapping = aes(x = month, y = Avo_sales_USD), color = "red") +
  facet_wrap(~region, ncol = 2) +
  ylab('USD')
```
As you can see from the plots, the relative rate at which the sales of avocados being sold in the US is increasing at a similar rate as that of housing costs. Thus, it is difficult to conclude whether Californian Millenials cannot afford to purchase houses due to their over consumption of avocados.


Work Cited:

https://stackoverflow.com/questions/65257153/rename-values-based-on-a-filter

https://stackoverflow.com/questions/33221425/how-do-i-group-my-date-variable-into-month-month-in-r
